
\begin{acknowledgements}      

\epigraph{If I have seen further it is by standing on the shoulders of Giants.}{\textit{Isaac Newton}}

I would like to express my deepest appreciation to my supervisor, Daniele Pucci, for believing in me and for giving me the opportunity to conduct my research in the \emph{Artificial and Mechanical Intelligence} laboratory. His guidance and advice have allowed me to grow more and more every day.

I am extremely grateful to Stefano Dafarra for all the time he spent discussing the research topics that are presented in this thesis. 

I could never reach the end of my Ph.D. without Silvio. I sincerely appreciate all of your time and work in advising me on several subjects. Every day, I am struck by your unwavering drive to assist all members of the community.

Olivier Stasse deserves special recognition for welcoming me to the Gepetto team in Toulouse. Thank you Olivier for giving me this wonderful opportunity. 

I cannot thank Nahuel enough. I thoroughly enjoyed every minute of our discussion.

My sincere gratitude goes to the reviewers, Marco Hutter and Jerry Pratt, for taking the time to read and review my thesis. I must thank Lorenzo Natale, Marco Maggiali, Ugo Pattacini, Vadim Tikhanoff, Luca Fiorio, and Alberto Parmigiani for their insightful comments during my Ph.D. research evaluations.

Words cannot explain how grateful I am to everyone on the \emph{Baciotti} team. Thank you, Giuseppe, Ines, Paolo, Riccardo, and Antonello for proving to be real friends. I never expected to meet folks like you.

A special thank goes to Prashanth! We embarked on this journey together and, unexpectedly, we survived. Thank you for all our chats! I have always admired your approach to research and you have always been an inspiration to me.

It is now Nicola's turn. Thank you for being my favorite housemate and adventure buddy. This thesis would not have been possible without your consistent encouragement to improve.

I would also like to thank all the guys and girls from LAAS, especially Gabriele, Gianluca, Fanny, and Côme. You made me feel at home even if I was abroad. 

Thanks should also go to current and past members of our lab: Yue, Carlotta, Claudia, Gianluca, Gabriele, Yeshasvi, Hosam, Lorenzo, Cisco, Punith, Fabio B., Diego, Kourosh, Raffaello, Gabriele, Italo, Affaf, Enrico, Tong, Milad, Anqing, Prajval, Venus, Valentino,  Ahmad and Mohamed, with whom I have
shared many unforgivable and great experiences.

I am sincerely thankful for the administrative support from Marta Caracalli, Lucia Betrò, and Valentina Scanarotti. 

Thank you for everything, Olivia, the sole genuine love of my life. Thank you for always being with me. Without you, I would be lost.

Special thanks also go to Stefano and Carmine, old friends who are living evidence that real friendship can transcend enormous distances.

Grazie Mamma, Babbo ed Eli. Non esistono parole per esprimere la mia gratitudine per tutto quello che avete fatto e che costantemente fate. Sarete per sempre una fonte di ispirazione che mi guiderà ogni giorno per il resto della mia vita. \\

\begin{flushright}
Giulio
\end{flushright}


\end{acknowledgements}
